### 2017-09-30 tex/article.tex
Initialised tex document

% !TEX root = ./article.tex

\documentclass{article}

\usepackage{mystyle}
\usepackage{myvars}

\begin{document}

  \maketitle

  \section{Ejercicio [TODO elegir uno] }

    \paragraph{}
    [TODO]

  \section{Ejercicio Acuicultura}

    \paragraph{}
    [TODO]

\end{document}